---
title: "Lab report"
author: "Yueqi"
date: "7/04/2021"
output: 
    html_document:
        toc: true
        highlight: default
---


<!--- Set global options that apply to every code chunk in this file, can be overwritten in individual chunk headers --->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

<!--- (Code Chunk 2) Load packages (previously installed)  --->
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

<!--- (Code Chunk 3) Set WD and read data file --->
```{r, include=FALSE}
# import data, fix NA and convert to numbers, checking for data type
rv <- read.csv("RVGE01.csv",na.strings=c(" NULL","","NA"))

library("readr")
rv$VE_any = parse_number(rv$VE_any)
rv$VE_severe = parse_number(rv$VE_severe)
rv$VE_Hosp = parse_number(rv$VE_Hosp)
str(rv)
rv$Prev.HIV = as.numeric(rv$Prev.HIV)

```

```{r}
str(rv)
```

subset to efficacy analysis
```{r}
rv_efficacy <- rv %>% filter(EffiX == 1)
dim(rv_efficacy)
```


subset to effectiveness analysis

```{r}
rv_effective <- rv %>% filter(EffiX == 0)
dim(rv_effective)

```



## Vaccine efficacy 
Concentrate on **Vaccine efficacy** against **severe RVGE**

### Multiple Linear Regression 
The outcome variable is **Vaccine Efficacy of Severe RVGE**
```{r}
# Effi_severe <- rv_efficacy %>% dplyr::select("VE_severe","Pop.Den","U5.Mor","Crude.BR","GDP","Safely.DW","Safely.San","Undernour","Prev.HIV" , "Gini.Index" , "Education" , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")
# 
# # evaluating the overall model
# summary(lm(VE_severe~ ., data = Effi_severe, na.action=na.omit))

# eliminate sanitation,prevalence of HIV and Education, as too many NAs would influence the result
Effi_severe02 <- rv_efficacy %>% dplyr::select("VE_severe","Pop.Den","U5.Mor","Crude.BR","GDP","Undernour" , "Gini.Index" , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")

### Here to scale the predictor variable
# Effi_severe02 = as.data.frame(scale(Effi_severe02[,-1]))
# Effi_severe02 = cbind(rv_efficacy$VE_severe,Effi_severe02)
# names(Effi_severe02)[1] <- "VE_severe"
# #finish scale
# 
summary(lm(VE_severe~ ., data = Effi_severe02))
```

#### AIC selection

* Forward selection
```{r}
library(MASS)

mod.full <- lm(VE_severe~ ., data = Effi_severe02)
mod.null <- lm(VE_severe~1, data = mod.full$model)
stepAIC(mod.null, scope = list(lower = mod.null, upper = mod.full), data = Effi_severe02, direction ='forward')

```

* Backward selection

```{r}

stepAIC(mod.full, scope = list(lower = mod.null, upper = mod.full), data = Effi_severe02, direction ='backward')

```

*Stepwise selection

```{r}

stepAIC(mod.null, scope = list(lower = mod.null, upper = mod.full), data = Effi_severe02, direction ='both')

```

### Spike and slab with `lm.spike`
```{r}
library(bsts)

spmodel <- lm.spike(VE_severe~ ., data = Effi_severe02, niter = 1000)
plot(spmodel,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"),burn = SuggestBurnLogLikelihood(spmodel$log.likelihood))

# remove predictor variable
spmodel1 <- lm.spike(VE_severe~  Pop.Den+Crude.BR +GDP + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop,     # remove Under 5 mortality
                   data = Effi_severe02, niter = 1000)
plot(spmodel1,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help") ,burn = SuggestBurnLogLikelihood(spmodel1$log.likelihood))

spmodel2 <- lm.spike(VE_severe~  Pop.Den+Crude.BR +GDP + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20,     # remove Under 5 mortality, urban population
                   data = Effi_severe02, niter = 1000)
plot(spmodel2,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"), burn = SuggestBurnLogLikelihood(spmodel2$log.likelihood))


```

### Collinearity between variables

```{r}
subset = Effi_severe02[,c(3,5,7,8,9,10)]   # select U5 mortality, GDP and Urban population
C = cor(subset)
C
library(corrplot)
corrplot(C, method="color")
```


### Sensitive analysis

```{r}
summary(lm(VE_severe~ U5.Mor + Urban.Pop, data = Effi_severe02))
```




#### Vaccine efficacy spike and slab

```{r}
# library(bsts)     # load the bsts package
# 
# ss <- AddLocalLevel(list(), y = Effi_severe02$VE_severe)
# 
# bsts_spike <- bsts(VE_severe ~ .,
#                state.specification = ss,
#                niter = 1000,
#                data = Effi_severe02,
#                expected.model.size = 5)  # Passed to SpikeSlabPrior.
# 
# plot(bsts_spike, "coef")


```

#### Remove some predictors and re-do the spike-and-slab

```{r}

# bsts_spike <- bsts(VE_severe ~ Pop.Den+Crude.BR +GDP + Undernour + Gini.Index
#                    + Extreme.Poverty+ Income.Low20+ Urban.Pop,     # remove Under 5 mortality
#                state.specification = ss,
#                niter = 1000,
#                data = Effi_severe02,
#                expected.model.size = 5)  # Passed to SpikeSlabPrior.
# 
# plot(bsts_spike, "coef")
# 
# bsts_spike <- bsts(VE_severe ~ Pop.Den+Crude.BR +GDP + Undernour + Gini.Index
#                    + Income.Low20+ Urban.Pop,     # remove Under 5 mortality, Extreme poverty
#                state.specification = ss,
#                niter = 1000,
#                data = Effi_severe02,
#                expected.model.size = 5)  # Passed to SpikeSlabPrior.
# 
# plot(bsts_spike, "coef")
# 
# bsts_spike <- bsts(VE_severe ~ Pop.Den+Crude.BR +GDP + Undernour + Gini.Index
#                    + Income.Low20,     # remove Under 5 mortality, Extreme poverty, Urban population
#                state.specification = ss,
#                niter = 1000,
#                data = Effi_severe02,
#                expected.model.size = 5)  # Passed to SpikeSlabPrior.
# 
# plot(bsts_spike, "coef")
# 
# bsts_spike <- bsts(VE_severe ~ Pop.Den+Crude.BR + Undernour + Gini.Index
#                    + Income.Low20,     # remove Under 5 mortality, Extreme poverty, Urban population, GDP
#                state.specification = ss,
#                niter = 1000,
#                data = Effi_severe02,
#                expected.model.size = 5)  # Passed to SpikeSlabPrior.
# 
# plot(bsts_spike, "coef")
# 
# # bsts_spike <- bsts(VE_severe ~ Pop.Den+ Undernour + Gini.Index
# #                    + Income.Low20,     # remove Under 5 mortality, Extreme poverty, Urban population, GDP, crude birthrate
# #                state.specification = ss,
# #                niter = 1000,
# #                data = Effi_severe02,
# #                expected.model.size = 5)  # Passed to SpikeSlabPrior.
# # 
# # plot(bsts_spike, "coef")
```



## Vaccine Effectiveness 
Concentrate on **Vaccine Effectiveness ** against **Hospitalized RVGE**

### Multiple Linear Regression 
The outcome variable is **Vaccine Effectiveness of Hospitalized RVGE**
```{r}
# Effect_hosp <- rv_effective %>% dplyr::select("VE_Hosp","Pop.Den","U5.Mor","Crude.BR","GDP","Safely.DW","Safely.San","Undernour","Prev.HIV" , "Gini.Index" , "Education" , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")
#  
# # evaluating the overall model 
# summary(lm(VE_Hosp~ ., data = Effect_hosp, na.action=na.omit))

# eliminate sanitation, too many NAs would influence the result
Effect_hosp02 <- rv_effective %>% dplyr::select("VE_Hosp","Pop.Den","U5.Mor","Crude.BR","GDP","Undernour", "Gini.Index"  , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")

# ### This part is for scale the data
# Effect_hosp02 = as.data.frame(scale(Effect_hosp02[,-1]))
# Effect_hosp02 = cbind(rv_effective$VE_Hosp,Effect_hosp02)
# names(Effect_hosp02)[1] <- "VE_Hosp"
# ### finish scale
# 
summary(lm(VE_Hosp~ ., data = Effect_hosp02))
```

* Forward selection 
```{r}
library(MASS)

# Full model (contains all predictors under consideration)
mod.fulleffective <- lm(VE_Hosp~ ., data = Effect_hosp02)
# Null model (intercept only, notice data= here)
mod.nulleffective <- lm(VE_Hosp ~1, data = mod.fulleffective$model)

stepAIC(mod.nulleffective, scope = list(lower = mod.nulleffective, upper = mod.fulleffective), data = Effect_hosp02, direction ='forward')

```

* Backward
```{r}
stepAIC(mod.fulleffective, scope = list(lower = mod.nulleffective, upper = mod.fulleffective), data = Effect_hosp02, direction ='backward')
```

* Stepwise

```{r}
stepAIC(mod.nulleffective, scope = list(lower =  mod.nulleffective, upper = mod.fulleffective), data = Effect_hosp02, direction ='both')
```

### Spike and slab with `lm.spike`
```{r}
library(bsts)

spmodel1 <- lm.spike(VE_Hosp~ ., data = Effect_hosp02, niter = 1000)
plot(spmodel1,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"))

spmodel2 <- lm.spike(VE_Hosp~ Pop.Den+Crude.BR + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop+ U5.Mor, data = Effect_hosp02, niter = 1000)    # remove GDP
plot(spmodel2,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"))

```

### Collinearity
```{r}
subset = Effect_hosp02[,c(2,3,5,10)]   # select U5 mortality, GDP and Urban population
C = cor(subset)
C
library(corrplot)
corrplot(C, method="color")
```


### Random forest
```{r}
library(randomForest)
m1 <- randomForest(
  formula = VE_severe~ .,
  data    = Effi_severe02,
  ntree=100, mtry=2, importance=TRUE)

m1
varImpPlot(m1)
```


```{r}
library(randomForest)
library(resample)      # data splitting 
library(randomForest) # basic implementation

Effect_hosp02 = na.omit(Effect_hosp02)
dim(Effect_hosp02)
m1 <- randomForest(
  formula = VE_Hosp~ .,
  data    = Effect_hosp02,
  ntree=500, mtry=2, importance=TRUE)

m1
varImpPlot(m1)
```

### Lasso regression
```{r}
x <- model.matrix(VE_severe ~ 0 + ., data = Effi_severe02) 
y <- Effi_severe02$VE_severe
library(glmnet)

#perform k-fold cross-validation to find optimal lambda value
lassoMod<- cv.glmnet(x, y, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- lassoMod$lambda.min
best_lambda

#produce plot of test MSE by lambda value
plot(lassoMod)

best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```


