---
title: "RVGE model_with revised spreadsheet data_0723"
author: "Yueqi"
date: "7/23/2021"
output: 
    html_document:
        toc: true
        highlight: default
---


<!--- Set global options that apply to every code chunk in this file, can be overwritten in individual chunk headers --->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

<!--- (Code Chunk 2) Load packages (previously installed)  --->
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

<!--- (Code Chunk 3) Set WD and read data file --->
```{r, include=FALSE}
# import data, fix NA and convert to numbers, checking for data type
rv <- read.csv("RVGE_0723_VECommunityControl.csv",na.strings=c("NULL",""))

rv <- rv %>% filter(GBD.Super.Region !="NA")

#subset to fewer columns
rv <- subset(rv[,c(3,4,11,13,15,18,51,52,53,54,55,56,57,58,59,60,61,62,63)])
```

```{r}
# str(rv)
rv$X..of.population.living.in.extreme.poverty....2.day. = as.numeric(rv$X..of.population.living.in.extreme.poverty....2.day.)
# check for the unique country number 
# unique(rv$Country)      

##replace the discontinued country name with the same one
rv$Country[rv$Country == c("United States","US","US Indian children")] = "US"

#After that, 52 country is included in our analysis
```

subset to efficacy/effectiveness analysis
```{r}
rv_efficacy <- rv %>% filter(Efficacy...1..or.effectiveness...0. ==1)
dim(rv_efficacy)
unique(rv_efficacy$Country)     # 38 countries is included in efficacy analysis

rv_effectiveness <- rv %>% filter(Efficacy...1..or.effectiveness...0. ==0)
unique(rv_effectiveness$Country)   # 31 countries is included in effectiveness analysis

```

## Vaccine efficacy 
Concentrate on **Vaccine efficacy** against **severe RVGE**

### Multiple Linear Regression 
The outcome variable is **Vaccine Efficacy of Severe RVGE**

```{r}
Effi_severe <- rv_efficacy[,-c(1,3,4,6)]


names(Effi_severe)[1:15] <- c("SuperRegion","VE_severe","Pop.Den","U5.Mor","Crude.BR","GDP","Safely.DW","Safely.San","Undernour","Prev.HIV" , "Gini.Index" , "Education" , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")

#perform multiple linear regression

Effi_severe <- Effi_severe[,-c(7,8,10,12)]    # As the water, education and HIV includes too many NA, excluded from our analysis
summary(lm(VE_severe~ ., data = Effi_severe, na.action=na.omit))
```


#### AIC selection
```{r}
library(MASS)

mod.full <- lm(VE_severe~ ., data = Effi_severe)
mod.null <- lm(VE_severe~1, data = mod.full$model)
stepAIC(mod.null, scope = list(lower = mod.null, upper = mod.full), data = Effi_severe, direction ='both')

```


### Spike and slab with `lm.spike`
```{r}
library(bsts)

spmodel <- lm.spike(VE_severe~ ., data = Effi_severe, niter = 500)
plot(spmodel,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"),burn = SuggestBurnLogLikelihood(spmodel$log.likelihood))

# remove predictor variable
# spmodel1 <- lm.spike(VE_severe~ SuperRegion+ Pop.Den+Crude.BR +GDP + Undernour + Gini.Index
#                    + Extreme.Poverty+ Income.Low20+ Urban.Pop,     # remove Under 5 mortality
#                    data = Effi_severe, niter = 1000)
# plot(spmodel1,
#      y = c("inclusion", "coefficients", "scaled.coefficients",
#               "residuals", "fit", "size", "help") ,burn = SuggestBurnLogLikelihood(spmodel1$log.likelihood))


```


## Random Forest for variable selection
```{r}
library(randomForest)
Effi_severe = na.omit(Effi_severe)
m1 <- randomForest(
  formula = VE_severe~ .,
  data    = Effi_severe,
  ntree=1000, mtry=2, importance=TRUE)

m1
varImpPlot(m1)
```


### Include the selected predictors in the MLR
```{r}
summary(lm(VE_severe~ SuperRegion+ U5.Mor + GDP+ Extreme.Poverty, data = Effi_severe, na.action=na.omit))

summary(lm(VE_severe~ SuperRegion+ U5.Mor, data = Effi_severe, na.action=na.omit))

summary(glm(VE_severe~ SuperRegion+ U5.Mor, data = Effi_severe, na.action=na.omit))
```

### Cross validation
```{r}
library(caret)
# Define training control
# train.control <- trainControl(method = "cv", number = 5)
train.control <- trainControl(method = "repeatedcv", 
                              number = 10, repeats = 50)
# Train the model
model1 <- train(VE_severe~ U5.Mor, data = Effi_severe, method = "lm",
               trControl = train.control)
model2 <- train(VE_severe~ SuperRegion+ U5.Mor, data = Effi_severe, method = "lm",
               trControl = train.control) 

model3 <- train(VE_severe~ SuperRegion+ U5.Mor + GDP , data = Effi_severe, method = "lm",
               trControl = train.control)

model4 <- train(VE_severe~ SuperRegion+ U5.Mor + GDP + Pop.Den , data = Effi_severe, method = "lm",
               trControl = train.control)

modelall <- train(VE_severe~ SuperRegion+ U5.Mor + GDP + Pop.Den+Crude.BR  + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop, data = Effi_severe, method = "lm",
               trControl = train.control)


print(model1)
print(model2)
print(model3)
print(model4)
print(modelall)


# Define training control: 5 fold cross-validation. If you want to perform 10 fold cv, set number=10,
# train_control <- trainControl(method="cv", number=5)
train.control <- trainControl(method = "cv", 
                              number = 7)
# Train the model using randomForest (rf)
# model <- train(Sepal.Length~., data=iris, trControl=train_control, method="rf")

model <- train(VE_severe~ SuperRegion+ U5.Mor, data = Effi_severe, method = "rf",
               trControl = train.control)
##The printed summary shows the sample sizes used, the best model selected and other information.
print(model)

train.control <- trainControl(method = "cv", 
                              number = 7)
modeltry <- train(VE_severe~ SuperRegion+ U5.Mor + GDP + Pop.Den+Crude.BR  + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop, data = Effi_severe, method = "rf",
               trControl = train.control)
print(modeltry)

# Make predictions
predictions <- predict(modelm, Effi_severe[,-1]) # Here the code actually not working... But finally the other form finally works!!! Congradulations!

# Summarize results
result <- data.frame(Actual=Effi_severe[,1],Predicted=predictions)
result$Difference <- abs(result$Actual – result$Predicted)
summary(result$Difference)



##===================## Final try, hope it would work...
dim(Effi_severe)
train_control<- trainControl(method="cv", number=10, savePredictions = TRUE)
model<- train(VE_severe~ SuperRegion+ U5.Mor, data = Effi_severe, trControl=train_control, method="rpart")
model$pred
print(model)

#=============#
train_control<- trainControl(method="cv", number=10, savePredictions = TRUE)
model<- train(VE_severe~ SuperRegion+ U5.Mor + GDP + Pop.Den+Crude.BR  + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop, data = Effi_severe, method = "rf", trControl=train_control)
model$pred
print(model)

#================#
# score = list()
# 
# LOOCV_function = function(x,label){
#  for(i in 1:nrow(x)){
#  training = x[-i,]
#  model = VE_severe~ SuperRegion+ U5.Mor
#  validation = x[i,]
#  pred = predict(model, validation[,setdiff(names(validation),label)])
#  score[[i]] = rmse(pred, validation[label]) # score/error of ith fold
#  }
#  return(unlist(score)) # returns a vector
 }
```

```{r}
# use random forest and repeated cv
library(caret)
# Define training control

train.control <- trainControl(method = "repeatedcv",
                              number = 10, repeats = 10,savePredictions = TRUE)
modelall <- train(VE_severe~ SuperRegion+ U5.Mor + GDP + Pop.Den+Crude.BR  + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop, data = Effi_severe, method = "rf",
               trControl = train.control)
modelall$pred
print(modelall)

modelmtry <- train(VE_severe~ SuperRegion+ U5.Mor , data = Effi_severe, method = "rf",
               trControl = train.control)
modelmtry$pred

print(modelmtry)

```

```{r}
# use linear regression and repeated cv
library(caret)
# Define training control

train.control <- trainControl(method = "repeatedcv",
                              number = 10, repeats = 10,savePredictions = TRUE)
modelall <- train(VE_severe~ SuperRegion+ U5.Mor + GDP + Pop.Den+Crude.BR  + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop, data = Effi_severe, method = "lm",
               trControl = train.control)
modelall$pred
print(modelall)

modelmtry <- train(VE_severe~ SuperRegion+ U5.Mor , data = Effi_severe, method = "lm",
               trControl = train.control)
modelmtry$pred

print(modelmtry)

```




###Vaccine Effectiveness

```{r}
Effect_hosp <- rv_effectiveness[,-c(1,3,4,5)]


names(Effect_hosp)[1:15] <- c("SuperRegion","VE_hosp","Pop.Den","U5.Mor","Crude.BR","GDP","Safely.DW","Safely.San","Undernour","Prev.HIV" , "Gini.Index" , "Education" , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")

#perform multiple linear regression

Effect_hosp <- Effect_hosp[,-c(7,8,10,12)]    # As the water, education and HIV includes too many NA, excluded from our analysis
summary(lm(VE_hosp~ ., data = Effect_hosp, na.action=na.omit))
```

#### AIC selection
```{r}
library(MASS)

mod.full <- lm(VE_hosp~ ., data = Effect_hosp)
mod.null <- lm(VE_hosp~1, data = mod.full$model)
stepAIC(mod.null, scope = list(lower = mod.null, upper = mod.full), data = Effect_hosp, direction ='both')

```


### Spike and slab with `lm.spike`
```{r}
library(bsts)

spmodel <- lm.spike(VE_hosp~ ., data = Effect_hosp, niter = 500)
plot(spmodel,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"),burn = SuggestBurnLogLikelihood(spmodel$log.likelihood))

# remove predictor variable
spmodel1 <- lm.spike(VE_hosp~ SuperRegion+ Pop.Den+Crude.BR + U5.Mor + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop,     # GDP
                   data = Effect_hosp, niter = 500)
plot(spmodel1,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help") ,burn = SuggestBurnLogLikelihood(spmodel1$log.likelihood))

```
```{r}
## Random Forest for variable selection
library(randomForest)
Effect_hosp = na.omit(Effect_hosp)
m2 <- randomForest(
  formula = VE_hosp~ .,
  data    = Effect_hosp,
  ntree=1000, mtry=2, importance=TRUE)

varImpPlot(m2)
```

### Cross validation
```{r}
library(caret)
# Define training control
# set.seed(123) 
train.control <- trainControl(method = "repeatedcv", 
                              number = 5, repeats = 50)
# Train the model
model1 <- train(VE_hosp~ GDP , data = Effect_hosp, method = "lm",
               trControl = train.control)
model2 <- train(VE_hosp~ GDP + U5.Mor, data = Effect_hosp, method = "lm",
               trControl = train.control)  
modelall <- train(VE_hosp~ GDP + U5.Mor+SuperRegion+ Pop.Den+Crude.BR + Undernour + Gini.Index + Extreme.Poverty+ Income.Low20 + Urban.Pop, data = Effect_hosp, method = "lm",trControl = train.control)


print(model1)
print(model2)
print(modelall)
```

```{r}
# use random forest and repeated cv
library(caret)
# Define training control

train.control <- trainControl(method = "repeatedcv",
                              number = 5, repeats = 10,savePredictions = TRUE)
modelall <- train(VE_hosp~ GDP + U5.Mor+SuperRegion+ Pop.Den+Crude.BR + Undernour + Gini.Index + Extreme.Poverty+ Income.Low20 + Urban.Pop, data = Effect_hosp, method = "rf",
               trControl = train.control)
modelall$pred
print(modelall)

modelmtry <- train(VE_hosp~ GDP + U5.Mor , data = Effect_hosp, method = "rf",
               trControl = train.control)
modelmtry$pred

print(modelmtry)
```

```{r}
#vaccine effectiveness
# use random forest and repeated cv
library(caret)
# Define training control

train.control <- trainControl(method = "repeatedcv",
                              number = 5, repeats = 10,savePredictions = TRUE)
modelall <- train(VE_hosp~ GDP + U5.Mor+SuperRegion+ Pop.Den+Crude.BR + Undernour + Gini.Index + Extreme.Poverty+ Income.Low20 + Urban.Pop, data = Effect_hosp, method = "lm",
               trControl = train.control)
modelall$pred
print(modelall)

modelmtry <- train(VE_hosp~ GDP + U5.Mor , data = Effect_hosp, method = "lm",
               trControl = train.control)
modelmtry$pred

print(modelmtry)
```


### Pool the effectiveness and efficacy data together
```{r}

Effect_hosp02 <- Effect_hosp %>% rename(VE =VE_hosp)
Effi_severe02 <- Effi_severe %>% rename(VE = VE_severe)

VE = rbind(Effi_severe02,Effect_hosp02)
```

```{r}
#MLR
summary(lm(VE~ ., data = VE, na.action=na.omit))

#Spike and slab
library(bsts)

spmodel <- lm.spike(VE~ ., data = VE, niter = 500)
plot(spmodel,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"),burn = SuggestBurnLogLikelihood(spmodel$log.likelihood))

#Random Forest
library(randomForest)
VE = na.omit(VE)
m2 <- randomForest(
  formula = VE~ .,
  data    = VE,
  ntree=1000, mtry=2, importance=TRUE)

varImpPlot(m2)

# use random forest and repeated cv
library(caret)
# Define training control
train.control <- trainControl(method = "repeatedcv",
                              number = 10, repeats = 10,savePredictions = TRUE)
modelall <- train(VE~ GDP + U5.Mor+SuperRegion+ Pop.Den+Crude.BR + Undernour + Gini.Index + Extreme.Poverty+ Income.Low20 + Urban.Pop, data = VE, method = "lm",
               trControl = train.control)
modelall$pred
print(modelall)

modelmtry <- train(VE~ GDP + U5.Mor , data = VE, method = "lm",
               trControl = train.control)
modelmtry$pred

print(modelmtry)



```






#######============################
split for the old coding part, should be discarded!!!

## Vaccine Effectiveness 
Concentrate on **Vaccine Effectiveness ** against **Hospitalized RVGE**

### Multiple Linear Regression 
The outcome variable is **Vaccine Effectiveness of Hospitalized RVGE**
```{r}
# Effect_hosp <- rv_effective %>% dplyr::select("VE_Hosp","Pop.Den","U5.Mor","Crude.BR","GDP","Safely.DW","Safely.San","Undernour","Prev.HIV" , "Gini.Index" , "Education" , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")
#  
# # evaluating the overall model 
# summary(lm(VE_Hosp~ ., data = Effect_hosp, na.action=na.omit))

# eliminate sanitation, too many NAs would influence the result
Effect_hosp02 <- rv_effective %>% dplyr::select("VE_Hosp","Pop.Den","U5.Mor","Crude.BR","GDP","Undernour", "Gini.Index"  , "Extreme.Poverty" ,"Income.Low20" , "Urban.Pop")

### This part is for scale the data
Effect_hosp02 = as.data.frame(scale(Effect_hosp02[,-1]))
Effect_hosp02 = cbind(rv_effective$VE_Hosp,Effect_hosp02)
names(Effect_hosp02)[1] <- "VE_Hosp"
### finish scale

summary(lm(VE_Hosp~ ., data = Effect_hosp02))
```

* Forward selection 
```{r}
library(MASS)

# Full model (contains all predictors under consideration)
mod.fulleffective <- lm(VE_Hosp~ ., data = Effect_hosp02)
# Null model (intercept only, notice data= here)
mod.nulleffective <- lm(VE_Hosp ~1, data = mod.fulleffective$model)

stepAIC(mod.nulleffective, scope = list(lower = mod.nulleffective, upper = mod.fulleffective), data = Effect_hosp02, direction ='forward')

```

* Backward
```{r}
stepAIC(mod.fulleffective, scope = list(lower = mod.nulleffective, upper = mod.fulleffective), data = Effect_hosp02, direction ='backward')
```

* Stepwise

```{r}
stepAIC(mod.nulleffective, scope = list(lower =  mod.nulleffective, upper = mod.fulleffective), data = Effect_hosp02, direction ='both')
```

### Spike and slab with `lm.spike`
```{r}
library(bsts)

spmodel1 <- lm.spike(VE_Hosp~ ., data = Effect_hosp02, niter = 1000)
plot(spmodel1,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"))

spmodel2 <- lm.spike(VE_Hosp~ Pop.Den+Crude.BR + Undernour + Gini.Index
                   + Extreme.Poverty+ Income.Low20+ Urban.Pop+ U5.Mor, data = Effect_hosp02, niter = 1000)    # remove GDP
plot(spmodel2,
     y = c("inclusion", "coefficients", "scaled.coefficients",
              "residuals", "fit", "size", "help"))

```

### Collinearity
```{r}
subset = Effect_hosp02[,c(2,3,5,10)]   # select U5 mortality, GDP and Urban population
cor(subset)

```







### Lasso regression
```{r}
#Lasso regression against vaccine efficacy
x <- model.matrix(VE_severe ~ 0 + ., data = Effi_severe02) 
y <- Effi_severe02$VE_severe
library(glmnet)

#perform k-fold cross-validation to find optimal lambda value
lassoMod <- cv.glmnet(x, y, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- lassoMod$lambda.min
best_lambda

#produce plot of test MSE by lambda value
plot(lassoMod)

best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```

### Lasso regression
```{r}
#Lasso regression against vaccine effectiveness
x <- model.matrix(VE_Hosp ~ 0 + ., data = Effect_hosp02) 
y <- Effect_hosp02$VE_Hosp
library(glmnet)

#perform k-fold cross-validation to find optimal lambda value
lassoMod <- cv.glmnet(x, y, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- lassoMod$lambda.min
best_lambda

#produce plot of test MSE by lambda value
plot(lassoMod)

best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```






```{r}
library(ggplot2)
library(sf)
library("dplyr")   
library("rnaturalearth")
library("rnaturalearthdata")
library(rgeos)

world <- ne_countries(scale = "medium", returnclass = "sf")

library("dplyr")
ve <- rv %>%
  filter(EffiX ==1 ) %>%
  select(Country, VE_severe) %>%                      # Select the two columns of interest
  rename(region = Country, ve = VE_severe)
world <- world %>% rename(region = admin)

VE_map <- left_join(ve, world, by = "region")

ggplot(data = VE_map) +
    geom_sf(aes(fill = ve)) +
    scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```

```{r}

library("rworldmap")
library("rworldxtra")
world <- getMap(resolution = "high")

library("dplyr")
ve <- rv %>%
  filter(EffiX ==1 ) %>%
  select(Country, VE_severe) %>%                      # Select the two columns of interest
  rename(region = Country, ve = VE_severe)

# VE_map <- left_join(world,ve, by = "region")
ggplot(data = world) +
    geom_sf()
ggplot(, aes( x = long, y = lat, group = group )) +
  geom_polygon(aes(fill = ve))
```

### Try again, WHF the output???
```{r}

library(ggplot2)
library(maps)
world = map_data("world")

ve <- rv %>%
  filter(EffiX ==1 ) %>%
  select(Country, VE_severe) %>%                      # Select the two columns of interest
  rename(region = Country, ve = VE_severe)

VE_map <- merge(ve,world)

ggplot(data = VE_map) + 
             geom_polygon(aes(x = long, y = lat, color = VE_map$ve, group = group))

```

### GeoMap (Finally success!!)
```{r}
library(tidyverse)
library(sf)
library(rvest)
library(stringr)
library(scales)
library("dplyr")

map.world <- map_data('world')
rv.geo <- read.csv('RV_age_group.csv',na.strings=c(" NULL","","NA","#VALUE!"))


## vaccine efficacy against severe RVGE
df.ve <- rv.geo %>%
  filter(EffiX ==1 ) %>%
  select(Country, VE_severe) %>%                      # Select the two columns of interest
  rename(country = Country, ve = VE_severe)

anti_join(df.ve, map.world, by = c('country' = 'region'))

map.VE <- left_join( map.world, df.ve, by = c('region' = 'country')) 

ggplot(map.VE, aes( x = long, y = lat, group = group )) +
  geom_polygon(aes(fill = ve)) + 
  labs(fill = 'VE'
       ,title = 'Vaccine Efficacy'
       ,subtitle = 'against severe RVGE'
       ,x = NULL
       ,y = NULL)

## vaccine effectiveness against hospitalization RVGE
df.vehosp <- rv.geo %>%
  filter(EffiX ==0 ) %>%
  select(Country, VE_hosp) %>%                      # Select the two columns of interest
  rename(country = Country, ve = VE_hosp)

anti_join(df.vehosp, map.world, by = c('country' = 'region'))

map.VE2 <- left_join( map.world, df.vehosp, by = c('region' = 'country')) 

ggplot(map.VE2, aes( x = long, y = lat, group = group )) +
  geom_polygon(aes(fill = ve)) + 
  labs(fill = 'VE'
       ,title = 'Vaccine Effectiveness'
       ,subtitle = 'against RVGE Hospitalization'
       ,x = NULL
       ,y = NULL)



```



### Age group analysis

```{r}
rv_age <- read.csv("")
```





